---
title: "CA8_plot"
format: html
editor: visual
execute: 
  output: true
  message: false
  warning: false
  echo: false
editor_options: 
  chunk_output_type: console
---

Challenges: 0: How to arrange each plot?

1: Chose the best plots from many options(sum, mean, table, raster).

2: The legend of the index, so that I can collect them

3: There is replication between table and spatial data

# Load functions

```{r}
library(terra)
library(tidyverse)
library(data.table)
library(ggpubr)
library(patchwork)


design <- c(area(1,1,1,2),
            area(2,1,2,1),
            area(2,2,2,2))
  
design2 <- c(area(1,1,1,3),
            area(2,1,2,1),
            area(2,2,2,2),
            area(2,3,2,3))
  
# The below functions only work when it is connected by the symbpl %>%  
add_unit <- function(data){ 
  mutate(
    data,
    unit=case_when(
      var=="npp"~list(expression(NPP~gC~m^-2~yr^-1)),
      var=="pr"~list(expression(Precipitation~mm~yr^-1)),
      var=="tas"~list("temperature Celsius")))
}

change_ext <- function(data){ 
 mutate(data,ext_type=str_replace(ext_type,"_"," & \n"))
}




extmodify <- function(data){ 
  mutate(data,
    region=str_replace_all(region,
    c("CA"="Central Asia","north"="NCA","central"="CCA",
    "sw"="SWCA","se"="SECA")))|> 
  mutate(region=factor(region,
    levels=c("Central Asia","NCA","SECA","SWCA","CCA")))}


spatial_theme <- theme(
      legend.position="bottom",    
      legend.key.width=unit(1.3, "cm"),
      legend.key.height = unit(0.7,"cm"),
      legend.text = element_text(size=10),
      legend.title = element_text(size=13),
      strip.background = element_rect(fill="white"))



space_theme <-  theme(
  plot.margin = unit(c(0, 0, 0, 0), "cm"),
  axis.text = element_blank(), 
  axis.ticks = element_blank(),
  legend.margin=margin(t = 0, unit='cm'),
  legend.position="bottom",    
  legend.key.width=unit(0.5, "cm"),
  legend.key.height = unit(0.3,"cm"),
  legend.title = element_text(size=8),
  strip.background = element_rect(fill="white"),
  panel.spacing = unit(0, "cm"))


##read climate bound ----
climatebound <- vect("F:/CMIP6_raster/INPUT/climate+dem+lucc+npp+bound/bound.shp")
values(climatebound) <-   c("NCA","CCA","SWCA","SECA")
  #c("north","central","sw","se")

bound_sf <- climatebound |> sf::st_as_sf()

```

# Read data

```{r}

# Import spatial mean SD trend created by CA2.5.
load("rdata/ca_mean_sd_trend.RData")
load("rdata/ca_density_sum80.RData")
load("rdata/ca_gross.RData")
load("rdata/ca_gross_space_relationship.RData")
load("F:/CMIP6_table/radata/contri_gross_plot.RData")




```

# Basic_table

```{r}

modifydata <- function(data){ data |> 
  mutate(
    country=str_replace_all(
      country,
      c("all"="Central Asia","north"="NCA","central"="CCA",
        "sw"="SWCA","se"="SECA")))|> 
  mutate(
    country=factor(
    country,
    levels=c("Central Asia","NCA","SECA","SWCA","CCA")))}



trend_zone_data <- alltrendresult_plot |> 
  filter(index=="npp") %>%
  modifydata()


sd_zone_data <- IAV_SD_table_plot |> 
  filter(index=="npp") |> 
  mutate(IAV_sd=ifelse(index=="tas",IAV_sd*100,IAV_sd))|> 
  modifydata()


contri_data <- contri_sum_plot %>% 
  select(scenario,model,central:sw) %>%  
  pivot_longer(central:sw,names_to = "country") %>% 
  modifydata()


corplot_data <- nppcorp_plot |> 
  select(-c(p_npp_pr:p_npp_tas)) |> 
  pivot_longer(cor_npp_pr:cor_npp_tas,values_to = "cor")%>%  # Shows the percent 
  mutate(name=str_replace_all(name,c("cor_npp_pr"="P","cor_npp_tas"="T"))) %>%
  modifydata()


contricli_data <- contricli_sum_plot %>% 
  select(scenario,name,model,central:sw) %>%  
  pivot_longer(central:sw,names_to = "country") %>% 
#  mutate(value=value*100) %>%  # Shows the percent 
  mutate(name=str_replace_all(name,c("pr_fit"="P","tas_fit"="T",".resid"="residual"))) %>%
  modifydata()



prop_nodate_zone |> 
  summarise(value=mean(value),.by=c(scenario,ext_type,region)) |> 
  filter(region=="CA",ext_type=="rest") 
```

# Basic_spatial

```{r}
 #mean_trend_sd_space

basic_space <- mean_nodatemodel_frame    |> 
  mutate(type="mean")                    |> 
  bind_rows(trend_nomodel_frame |> mutate(type="trend")) |> 
  bind_rows(sd_nomodel_frame |> mutate(type="sd"))       |> 
#  filter(scenario%in%c("SSP585")) |> 
  select(-label)                                         |> 
  group_nest(type,var)                                   |> 
  mutate(bound=list(sf::st_as_sf(climatebound)))         |> 
  add_unit()                                             |> 
  mutate(
    unit=ifelse(
      var=="npp"&type=="trend",list(expression(NPP~gC~m^-2~yr^-2)),unit))
  


    
  
    
basic_space_plot <- pmap(
  list(basic_space$data,
       basic_space$bound,
       basic_space$unit),
  ~ggplot()+
    geom_raster(data =..1,aes(x = x, y = y,fill = mean)) +
    geom_sf(data=..2,fill=alpha(0.01)) +
    geom_sf_text(aes(label=value),size=4,data = ..2)    +
    coord_sf()+
  # scale_fill_gradientn(
  #   colours = c("#ffffcc", "#a1dab4", "#41b6c4", "#2c7fb8", "#253494"),
  #   name = expression("NPP gC m"^{-2})
  # )   +
   scale_fill_gradientn(
      colours = colorRampPalette(
        c("#ffffcc", "#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#0c2c84")
      )(200),  # ⬅️ more color levels (200 shades)
      name = expression("NPP gC m"^{-2})
    ) +
  
    labs(fill=..3, x ='',y="")+
    theme_bw()          +     
    spatial_theme+ 
    facet_wrap(.~scenario))

basic_space_plot[[4]]
 
```

# Sensitivity spatial

```{r}

relation_space <- bind_rows(
  model_out |>  pivot_longer(rsquare:pr_sen) ,
  cor_out   |>  pivot_longer(pr_npp:tas_npp_p) ,
  npp_contri |> select(-npp) |> pivot_longer(contri) |> mutate(value=10000*value),
  cli_contri |> rename(name=var)|> mutate(value=10000*value))  |> 
  filter(scenario=="SSP585") |> 
  group_nest(name) |> # There is NA value in dataset.
  mutate(
    mean_data=map(
      data, 
      ~summarise(.,value=mean(value,na.rm=T),.by = c(scenario,x,y))),
    sd_data=map(
      data, 
      ~summarise(.,value=sd(value,na.rm=T),.by = c(scenario,x,y)))) |> 
  select(-data)


relation_space_data <- 
  relation_space |> 
  mutate(bound=list(sf::st_as_sf(climatebound))) |> 
  mutate(name=as.list(name))


# With boundary but not detailed information in data frame 
relation_mean_plot <- 
  pmap(list(relation_space_data$mean_data,
            relation_space_data$bound,
            relation_space_data$name),
       ~ggplot()+
         geom_raster(data =..1,aes(x = x, y = y,fill = value)) +
         geom_sf(data=..2,fill=alpha(0.01)) +
         geom_sf_text(aes(label=value),size=5,data = ..2 ) +
         coord_sf()+
         scale_fill_gradientn(colours = terrain.colors(100))+
         labs(fill=..3, x ='',y="")+
         theme_bw() +
         spatial_theme) 

relation_mean_plot[[1]]
```

# Result0 extremes

```{r}
load("rdata/ca_density_sum80.RData")

ext_nodate_zone %>% 
  group_by(scenario,var,magnitude,index,region) %>% 
  summarise(value=mean(value)) %>% 
  filter(index=="value") %>% 
  filter(region%in%c("CA")) %>% View()


plot1 <- ext_nodate_zone %>% filter(!var=="npp")  %>% 
  extmodify

p5 <- plot1 %>%
  filter(index == "value") %>%
  ggplot(aes(x = region, y = value, fill = scenario)) +
  stat_summary(
    fun = mean,
    geom = "col",
    position = position_dodge(width = 0.9),
    width = 0.8
  ) +
  stat_summary(
    aes(color = scenario),
    fun.data = mean_sd,
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2,
    linewidth = 0.8
  ) +
  facet_grid(
    var ~ magnitude,
    scales = "free_y",
    switch = "y",
    labeller = labeller(
      var = as_labeller(c(pr = "Precipitation IAV (mm)", tas = "Temperature IAV (°C)")),
      magnitude = as_labeller(c(tenth = "Negative extreme", ninetieth = "Positive extreme"))
    )
  ) +
  theme_bw()+
  theme(
    legend.position = "top",
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(size = 12)
  )+
  labs(y="",x="",fill="",col="") 

# Save the plot
ggsave(
  filename = "paper_figure/climate_Extremes.png",  # you can change the file name and extension
                               # the ggplot object
  width = 8,                               # width in inches
  height = 6,                              # height in inches
  dpi = 300                                # resolution
)



# Base plotting function


# Function to generate a single plot
make_plot <- function(data, var_filter, mag_filter, fill_lab, bound) {
  data %>%
    filter(var == var_filter, magnitude == mag_filter) %>%
    ggplot(aes(x = x, y = y, fill = value)) +
    geom_tile() +
    geom_sf(data = bound, fill = NA, color = "black", inherit.aes = FALSE) +
    geom_sf_text(aes(label = value), size = 3, data = bound, inherit.aes = FALSE) +
    facet_wrap(~scenario) +
    scale_fill_gradientn(
      colours = c("#ffffcc", "#a1dab4", "#41b6c4", "#2c7fb8", "#253494")
    ) +
    labs(
      fill = fill_lab,
      x = "",
      y = "",
      title = paste0(
        ifelse(var_filter == "pr", "Precipitation", "Temperature"), " - ",
        ifelse(mag_filter == "tenth", "Negative IAV extreme", "Positive IAV extreme")
      )
    ) +
    theme_minimal() +
    theme(strip.placement = "outside",
          legend.position="bottom")
}


# Build the 4 plots
p1 <- make_plot(
  ext_nodatemodel,
  "pr",
  "tenth",
  "Precipitation (mm)",
  sf::st_as_sf(climatebound)
)

p2 <- make_plot(ext_nodatemodel, "pr",  "ninetieth", "Precipitation (mm)",  sf::st_as_sf(climatebound))
p3 <- make_plot(ext_nodatemodel, "tas", "tenth",     "Temperature (°C)",  sf::st_as_sf(climatebound))
p4 <- make_plot(ext_nodatemodel, "tas", "ninetieth", "Temperature (°C)",  sf::st_as_sf(climatebound))

ggsave("p1_precip_tenth.png",  plot = p1, width = 5, height = 4, dpi = 300)
ggsave("p2_precip_ninetieth.png", plot = p2, width = 5, height = 4, dpi = 300)
ggsave("p3_temp_tenth.png",    plot = p3, width = 5, height = 4, dpi = 300)
ggsave("p4_temp_ninetieth.png", plot = p4, width = 5, height = 4, dpi = 300)






grid_4 <- (p1 | p2) /
          (p3 | p4)


grid_4

ggsave("extremes_grid.png", 
       plot = final_plot, 
       width = 10, height = 10, dpi = 600)

   
```

# Result1 (NPP:Trend,SD,contribution)

```{r}
# Add significant line in the trend plot

list(alltrendresult_plot,
IAV_SD_table_plot,
contri_sum_plot)


sd_plot <- ggplot(
  sd_zone_data,aes(x=country,y=IAV_sd,fill=scenario))+
  stat_summary(
    fun = mean,geom = "bar",
    position=position_dodge(0.9),width = 0.9) +
  stat_summary(
    fun.min = function(x) {mean(x) - sd(x)/sqrt(length(x))},
    fun.max = function(x) {mean(x) + sd(x)/sqrt(length(x))},
    geom = "errorbar",
    position=position_dodge(0.9),width = 0.9) +
  labs(y=expression(Annual~NPP~variation~TgC~yr^-1),fill="",x="")+
  theme_bw()+
  theme(
         strip.text =  element_text(size=10,family = "sans"),
         axis.text.x = element_text(size=10,family = "sans"),
         axis.text.y = element_text(size=10,angle=90,family = "sans"),
         axis.ticks.length =  unit(-0.1, "cm"),
         legend.text=element_text(size=8,family = "sans"),
         legend.position = c(0.55,0.9),
         legend.direction = "horizontal") 
  

sd_plot


load("rdata/contri_year_gross.RData")

sum_lab <- as_labeller(c(
  npp = "NPP~TgC~yr^-1", 
  pr="Precipitation mm/yr",
  tas="Temperature Celsius/yr"),label_parsed)

trend_time_plot <- ggplot(allmeanerro_plot |> 
         filter(country=="all",index=="npp"),
       aes(x = year, y = mean, fill=Scenario))+
  geom_ribbon(aes(ymin = mean-erro,ymax =mean+erro),alpha = 0.2) +
  geom_line(aes(col=Scenario)) +
  geom_smooth(method = "lm",se=FALSE,aes(col=Scenario))+
  facet_wrap(.~index,ncol = 1,
             scales="free_y",
             strip.position = "left",
             labeller= sum_lab)+
  labs(y="",x="",fill="",col="") +
  theme_bw()+
  theme( strip.placement = "outside",
         strip.background = element_blank(),
         strip.text =  element_text(size=10,family = "sans"),
         axis.text.x = element_text(size=10,family = "sans"),
         axis.text.y = element_text(size=10,angle=90,family = "sans"),
         axis.ticks.length =  unit(-0.1, "cm"),
         legend.text=element_text(size=10,family = "sans"),
         legend.position = c(0.45,0.9),
         legend.direction = "horizontal") 


trend_time_plot




basic_space_plot2  <- basic_space_plot[[7]]+
   labs(title =  "Annual NPP Trend" )+
  basic_space_plot[[4]] +
  labs(title =  "Annual NPP Variation" )
   # relation_mean_plot[[1]]+
   # labs( title =  expression("Contribution to NPP IAV"),
   #       fill=expression("Contribution"~10^-4~"%")) 









basic_space_plot2
  



trend_time_plot+sd_plot+
  basic_space_plot2+
  plot_layout(design = c(area(1,1,1,1),  # up left down right
                       area(1,2,1,2.5),
                       area(2,1,2,2.5)))+

      plot_annotation(tag_levels = 'a')



ggsave("paper_figure/npp.png",height=6.5,width=8,dpi=700,scale=1.4)


basic_space_plot[[7]] +
annotation_custom(grob=ggplotGrob(trend_plot),
                     xmin=50,xmax=80,
                     ymin=35,ymax=45)


lineplot   <-  trend_time_plot +
  #trend_plot            +
  # sd_plot       + 
  # contri_plot   +
   plot_layout(guides="collect") &
   theme(legend.position='bottom'
        )
```

# New result1

```{r}
# Load density in table 
load("F:/CMIP6_table/radata/contri_density_plot.RData")






p1 <- basic_space_plot[[4]]+  labs(title =  "Annual NPP Variation" )


 
area <-   caarea_zone %>% 
    mutate(country=c("NCA","CCA","SWCA","SECA")) %>% 
    select(-c(value,unit))  %>% 
    rbind(  
  caarea_zone %>% 
  summarise(area=sum(area)) %>% 
  mutate(country="Central Asia"))
    
  
sd_zone_data2 <- sd_zone_data %>% 
  left_join(area) %>% 
  mutate(IAV_sd=IAV_sd*1e12/area*2)



sd_data3 <- sd_zone_data2 %>% 
  mutate(type="density") %>% 
  bind_rows(mutate(sd_zone_data,type="whole"))


sd_plot2 <- ggplot(
  sd_data3 |> 
  mutate(country=factor(country,
    levels=c("Central Asia","NCA","SECA","SWCA","CCA"))), 
  
  aes(x = country, y = IAV_sd, fill = scenario)
) +
  stat_summary(
    fun = mean, geom = "bar",
    position = position_dodge(0.9), width = 0.9
  ) +
  stat_summary(
    fun.min = function(x) { mean(x) - sd(x)/sqrt(length(x)) },
    fun.max = function(x) { mean(x) + sd(x)/sqrt(length(x)) },
    geom = "errorbar",
    position = position_dodge(0.9), width = 0.9
  ) +
  facet_grid(
    type ~ .,
    scales = "free_y",
    switch = "y",
    labeller = labeller(
      type = as_labeller(
        c(
          density = "Annual~NPP~variation~gC~m^{-2}~yr^{-1}",
          whole   = "Annual~NPP~variation~TgC~yr^{-1}"
        ),
        label_parsed
      )
    )
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    strip.placement = "outside",
    strip.background = element_blank(),              
    strip.text.y.left = element_text(size=11, family="sans"),
    axis.text.x = element_text(size=10, family="sans"),
    axis.text.y = element_text(size=10, angle=90, family="sans"),
    axis.ticks.length = unit(-0.1, "cm"),
    legend.text = element_text(size=8, family="sans"),
    legend.direction = "horizontal"
  ) +
  labs(y = NULL,fill= NULL,x=NULL) # remove redundant y-axis label

sd_plot2



ggsave(
  filename = "paper_figure/annual_npp_variation.png",
  plot = p1,
  width = 6, height = 5, dpi = 300,
  bg = "transparent"
)

# Plot 2: sd_plot
ggsave(
  filename = "paper_figure/sd_plot2.png",
  plot = sd_plot2,
  width = 5, height = 6, dpi = 300,
  bg = "transparent"
)


```

# Result2 (Correlation/Contribution/Sensitvity)

```{r}
load("rdata/ca_gross_space_relationship.RData")
load("rdata/ca_gross_space_relationship_extreme.RData")
load("rdata/ca_gross_space_relationship_extreme_meanmodel.RData")



```

## Contribution

```{r}
# Question 

# Normally

climatecon_space_data <- cli_contri|>
  rename(name=var)|>
  mutate(value=10000*value)  |>
  filter(scenario=="SSP585") |>
  summarise(value=mean(value,na.rm=T),.by = c(scenario,x,y,name)) |>
  group_nest(scenario)|>
  mutate(bound=list(sf::st_as_sf(climatebound)))  |>
  mutate(name=as.list("dd"))


hist(climatecon_space_data$data[[1]]$value)
range( climatecon_space_data$data[[1]]$value)
quantile(climatecon_space_data$data[[1]]$value)

8*3/11

color2 <- c("#d62e29","#e06345", "#fab884", 
            "#7bed01","#06d41a","#1ba87d","#b2c7d8",
           "#97aebe","#097da4","#0951a4","#2409a4")
 

label2 <-  c("-0.3", "-0.2", "-0.1",
             "0","0.1", "0.2","0.4",
             "0.8","1.6","3.2","4","6")

length(color2)
length(label2)


contribution_discrete <-   pmap(
  list(climatecon_space_data$data, climatecon_space_data$bound),
  ~ ggplot() +
  geom_raster(data = ..1,
              aes(x = x,y = y,fill = cut(value,label2)
              ))                                    +
  geom_sf(data=..2, fill = alpha(0.01))             +
  geom_sf_text(aes(label=value),size=5,data = ..2 ) +
  coord_sf()                                        +
  facet_grid(. ~ name,
             labeller = as_labeller(c("pr"="Precipitation",
                                      "tas"="Temperature")))+
  scale_fill_manual(
    na.value = "white",
    values =color2,
    labels = label2) +
    guides(
      fill = guide_legend(title.position="top",
        title = expression("Contribution to NPP IAV"~10^-4~"%"),
        label.position = "bottom",
        nrow = 1,
        keywidth = 2.18,
        keyheight=1)
      ) +
    labs(x = "", y = "") +
    theme_classic() +
    theme(
      strip.text = element_text(size=12),
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_blank(),
      legend.position = c(0.5,-0.2),
      legend.spacing.x = unit(0, 'cm'),
      legend.text.align = 0,
      plot.margin = margin(0,0,40,0)
    ))

contribution_discrete

```

## Correlation

```{r}
cor_p_data <- cor_out   |>  
  pivot_longer(pr_npp:tas_npp_p)|> 
  filter(scenario=="SSP585") |> 
  filter(str_detect(name,"_p"))  |> 
  summarise(value=mean(value,na.rm=T),.by = c(scenario,x,y,name))|>       
  mutate(p_value=ifelse(value<0.05,value,NA),.keep = "unused") |> 
  mutate(name=str_replace(name,"_p",""))

cor_data <- cor_out   |>  
  pivot_longer(pr_npp:tas_npp_p)|> 
  filter(scenario=="SSP585") |> 
  filter(!str_detect(name,"_p"))  |> 
  summarise(value=mean(value,na.rm=T),.by = c(scenario,x,y,name)) |> 
  left_join(cor_p_data)                   |> 
  group_nest(scenario)|> 
  mutate(bound=list(sf::st_as_sf(climatebound)))  



cor_data$data[[1]]$value |> quantile()
cor_data$data[[1]]$value |> hist()

cut(cor_data$data[[1]]$value,c(-0.3,-0.2,-0.1,
                               0,0.2,0.4,0.5,0.6,0.7))


color3 <- c("#d62e29","#e06345", "#fab884", 
            "#7bed01","#06d41a","#1ba87d",
            "#b2c7d8","#097da4")



length(color3)

label3 <- c("-0.3","-0.2","-0.1",
            "0","0.2", "0.4",
            "0.5","0.6","0.7")

length(label3)


correlation_discrete <- pmap(list(cor_data$data,
          cor_data$bound),
          ~ggplot()+
     geom_raster(data = ..1,
     aes(x = x, y = y,fill=cut(value,label3))) +
     geom_sf(data=..2,fill=alpha(0.01)) +
     geom_sf_text(aes(label=value),size=5,data = ..2 ) +
     geom_point(data=..1 |> filter(!is.na(p_value)),
               aes(x=x,y=y),shape=1,size=0.1,alpha=0.5)  +
    coord_sf()                                            +
    facet_grid(.~name,labeller = as_labeller(c("pr_npp"="Precipitation",
                                               "tas_npp"="Temperature")))   +
    scale_fill_manual(na.value = "white",
           values = color3,
           labels=label3)+ 
    guides(fill=guide_legend(title.position="top",
                             title="Correlation",
                             label.position = "bottom",nrow = 1,
                             keywidth=3,keyheight = 1)) +
    labs(x="",y="") +
    theme_classic() +
    theme(
      strip.text = element_text(size=12),
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_blank(),
    #  legend.position = "bottom",
          legend.position = c(0.55,-0.2), 
     legend.spacing.x = unit(0, 'cm'),
     legend.text.align = 0,
     plot.margin = margin(0,0,30,0)
    ))
```

## sensitivity

```{r}

sen_theme <- 
    theme(
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_blank(),
     legend.position = c(0.55,-0.2), 
     legend.spacing.x = unit(0, 'cm'),
     legend.text.align = 0,
     legend.key.width =unit(2.5,"cm"),
     legend.key.height =unit(1,"cm"),
     legend.title = element_text(size=9),
     plot.margin = margin(0,0,30,0)
    )

model_fit_p1 |> 
  filter(scenario=="SSP585") |>
  summarise(tas_p=mean(tas,na.rm=T),pr_p=mean(pr,na.rm=T),.by = c(scenario,x,y)) 

sensitivity_space_data <- model_out |> 
  left_join(model_fit_p1)           |> 
  select(-rsquare) |> 
  filter(scenario=="SSP585") |>
  summarise(across(where(is.numeric),mean),.by = c(scenario,x,y)) |> 
  mutate(pr_sen=pr_sen*100)

sensitivity_space_data |> drop_na(pr_sen,pr) |> filter(pr<0.05) 

sensitivity_space_data |> drop_na(tas_sen,tas) |> filter(tas<0.05)  

color4 <- c("#d62e29","#e06345", 
            "#7bed01","#06d41a",
            "#1ba87d")



pr_cut <- sensitivity_space_data$pr_sen   |> 
  quantile(probs = seq(0, 1, 0.2),na.rm=T) |> 
  as.numeric() |> 
  round(digits = 4)

pr_cut
pr_cut <- c(0,0.05,0.06,0.07,0.08,0.12)

ext_cut <- c(-0.1,0,0.1,0.15,0.2,0.3)

prsensitivity_plot<- ggplot()+
     geom_raster(data = sensitivity_space_data |> drop_na(pr_sen,pr),
     aes(x = x, y = y,fill=cut(pr_sen,
                               ext_cut)))+
                               #pr_cut))) +
     geom_sf(data=bound_sf,fill=alpha(0.01)) +
     geom_sf_text(aes(label=value),size=5,data =bound_sf) +
     geom_point(data=sensitivity_space_data |>
                  drop_na(pr_sen,pr) |> 
                  filter(pr<0.05),
                aes(x=x,y=y),shape=1,size=0.1,alpha=0.5)  +
    coord_sf()                                            +
    scale_fill_manual(na.value = "white",
           values = color4,
          # labels=pr_cut
           labels=ext_cut
           )+ 
    guides(fill=guide_legend(title.position="top",
                             title = expression(Sensitivity~of~NPP~to~precipitation(Tg~100~mm^{-1})), #100m
                             label.position = "bottom",nrow = 1,
                             keywidth=2,keyheight = 1)) +
    labs(x="",y="") +
    theme_classic() +
    sen_theme


prsensitivity_plot

tas_cut <- sensitivity_space_data$tas_sen |> 
  quantile(probs = seq(0, 1, 0.2)) |> 
  as.numeric(digits = 3)  |> 
  round(digits = 4)

tas_cut


tas_cut <- c(-0.02,-0.01,0,0.001,0.003,0.03)


tas_sensitivity_plot <-ggplot()+
     geom_raster(data = sensitivity_space_data,
     aes(x = x, y = y,fill=cut(tas_sen,tas_cut))) +
     geom_sf(data=bound_sf,fill=alpha(0.01)) +
     geom_sf_text(aes(label=value),size=5,data =bound_sf) +
     geom_point(data=sensitivity_space_data |> filter(tas<0.05),
                aes(x=x,y=y),shape=1,size=0.1,alpha=0.5)  +
    coord_sf()                                            +
    scale_fill_manual(na.value = "white",
           values = color4,
           labels=tas_cut)+ 
    guides(fill=guide_legend(title.position="top",
                             title = expression(Sensitivity~of~NPP~to~temperature~(Tg~degree*C^{-1})),
                             label.position = "bottom",nrow = 1,
                             keywidth=2,keyheight = 1)) +
    labs(x="",y="") +
    theme_classic() +
    sen_theme
 
tas_sensitivity_plot

```

## Variation

```{r}

variation_plot <-  basic_space |> 
       filter(!var=="npp",type%in%c("sd")) |> 
       select(-c(bound,unit)) |> 
       unnest(data)           |> 
       pivot_wider(names_from = var,values_from = mean)


quantile(variation_plot$pr) 
quantile(variation_plot$tas)



pr_var_cut <- c(30,50,70,100,200)
tas_var_cut <- c(0.5,0.6,0.8,1,1.2) 


color4 <- c("#06d41a","#1ba87d", 
            "#097da4","#2409a4")


variation_pr <- ggplot()+
     geom_raster(data = variation_plot,
     aes(x = x, y = y,fill=cut(pr,pr_var_cut))) +
     geom_sf(data=bound_sf,fill=alpha(0.01)) +
     geom_sf_text(aes(label=value),size=5,data =bound_sf) +
    coord_sf()                                            +
    scale_fill_manual(na.value = "white",
           values = color4,
           labels=pr_var_cut)+ 
    guides(fill=guide_legend(title.position="top",
                             title="The variation of precipitation (mm)", #100m
                             label.position = "bottom",nrow = 1,
                             keywidth=2,keyheight = 1)) +
    labs(x="",y="") +
    theme_classic() +
    sen_theme

variation_pr 


variation_tas <- ggplot()+
     geom_raster(data = variation_plot,
     aes(x = x, y = y,fill=cut(tas,tas_var_cut))) +
     geom_sf(data=bound_sf,fill=alpha(0.01)) +
     geom_sf_text(aes(label=value),size=5,data =bound_sf) +
    coord_sf()                                            +
    scale_fill_manual(na.value = "white",
           values = color4,
           labels=tas_var_cut)+ 
    guides(fill=guide_legend(title.position="top",
                             title="The variation of temperature (℃)", #100m
                             label.position = "bottom",nrow = 1,
                             keywidth=2,keyheight = 1)) +
    labs(x="",y="") +
    theme_classic() +
    sen_theme


variation_tas 


```

## Save plot (continuous method)

```{r}


contribution_discrete[[1]]/correlation_discrete[[1]]+ 
  (prsensitivity_plot+tas_sensitivity_plot)+ plot_annotation(tag_levels="a") 

ggsave("paper_figure/cor_contri_sen_dis.png",width=6.5,height=10,dpi=700,scale=1) 


(correlation_discrete[[1]]|contribution_discrete[[1]])/ 
  (prsensitivity_plot|tas_sensitivity_plot| 
  variation_pr|variation_tas)+ plot_annotation(tag_levels="a") 


# (correlation_discrete[[1]]|prsensitivity_plot|tas_sensitivity_plot)/ 
#   (contribution_discrete[[1]]|variation_pr|variation_tas)+ plot_annotation(tag_levels="a") 

ggsave("paper_figure/cor_contri_sen_variation.png",width=13,height=6,dpi=700,scale=1.05) 


contribution_discrete[[1]]/correlation_discrete[[1]]+ plot_annotation(tag_levels="a") 

ggsave("paper_figure/cor_contri_dis.png",width=6,height=6,dpi=700) 

prsensitivity_plot+tas_sensitivity_plot+ plot_annotation(tag_levels="a") 

ggsave("paper_figure/sensitivy.png",width=7,height=3.5,dpi=700,scale=1) 


  relation_mean_plot[[3]]+
#  annotate(geom="text",label= "Precipitation",x=52,y=54,size=4)+
  labs(fill="Correlations",title = "Precipitation") +
  relation_mean_plot[[8]]+
  labs(fill="Correlations",title = "Temperature")+  
  contribution_plot    +
  plot_annotation(tag_levels = "a")+
  plot_layout(design = c(area(1,1,1,1),  # up left down right
                       area(1,2,1,2),
                       area(2,1,2,2)))

  # design = c(area(1,1,1,1),
  #            area(1,2,1,2)))   &
  # theme(legend.position='top')
  
ggsave("paper_figure/cor_contri.png",width=6,height=6,dpi=700,scale=1.2) 


relation_mean_plot[[6]]


ggsave("paper_figure/r_square.png",width=3,height=3,dpi=700,scale=1.5) 

relation_mean_plot[[5]]
relation_mean_plot[[10]]
```

# Result3 Extreme attribution Proportion

```{r}
#| fig-height: 6
#| fig-width:  8
#| fig-dpi: 700




prop_nodate_zone %>% 

  group_by(ext_type,magnitude,region,scenario) %>% 
  summarise(value=mean(value)) %>% 
  filter(region=="se") %>% 
  filter(ext_type%in%c("dry","wet"))


plot_prop_nodate_zone <- prop_nodate_zone |>
    mutate(ext_type = str_replace(ext_type, "_", " & \n")) |> 
    filter(scenario == "SSP585") |> 
    mutate(magnitude=str_replace_all(magnitude,c("ninetieth"="positive","tenth"="negative"))) %>% 
    extmodify()


plot_abc <- plot_prop_nodate_zone %>% 
    filter(!ext_type%in%c("rest","hot & \nnotdry"))

plot_abc$ext_type <- factor(plot_abc$ext_type,levels=c(
  "dry", "wet", "cold", "hot",
  "dry & \nhot", "dry & \nnothot"
)
)




prop_zone_plot <- ggplot(
  data = plot_abc,
  mapping = aes(
    x = ext_type,
    y = value,
    fill = region
  )
) +
  geom_boxplot(
    width = 0.6,
    outlier.size = 1,
    outlier.alpha = 0.6
  ) +
  scale_y_continuous(
    limits = c(0, 0.75),
    expand = c(0, 0),
    breaks = seq(0, 0.75, by = 0.2),
    labels = function(x) ifelse(x == 0, "", scales::percent_format(accuracy = 1)(x))
  ) +
  facet_wrap(
    ~ magnitude,
    ncol = 1,
    strip.position = "left"
  ) +
  labs(
    fill = "",
    x = "",
    y = "Fraction of NPP Extremes Attributed to Climate Extremes"
  ) +
  theme_bw() +  # Move theme_bw BEFORE theme()
  theme(
    axis.text.x = element_text( hjust = 1, vjust = 1),
    axis.ticks.length.y = unit(-0.1, "cm"),
    axis.text.y = element_text(angle = 90),
    strip.background = element_blank(),
    strip.placement = "outside",
    legend.position = "top",
    legend.box.spacing = unit(0.5, "cm"),
    strip.text = element_text(size = rel(1.1)),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    plot.tag = element_text(size = 14, face = "bold")  # Make sure tag is visible
  ) 


prop_zone_plot

ggsave(
  filename = "paper_figure/fraction_attributed_events_region.tiff",
  plot = prop_zone_plot,
  width = 8,
  height = 6,
  scale= 0.9,
  units = "in",
  dpi = 600,
)












# Combine this with the plot produced in the whole in PPT.

prop_zone_plot <- ggplot(
  prop_nodate_zone |>
    filter(value<1) |> 
    mutate(ext_type=str_replace(ext_type,"_"," & \n")) |> 
   # filter(scenario=="SSP585")  |> 
    filter(region=="CA")  |> 
    extmodify(),
  #aes(x=ext_type,y=value,fill=region))+
  aes(x=ext_type,y=value,fill=class))+
  #aes(x=ext_type,y=value,fill=scenario))+
  geom_boxplot()+
  scale_y_continuous()+
  labs(title = "",x="",fill="",y="Fraction of attributed events")+
  theme_bw() +
  theme(axis.ticks.length.y = unit(-0.1,"cm"),
        axis.text.y = element_text(angle=90),
        strip.background = element_blank(),
        strip.placement = "outside")+
#  facet_wrap(class~magnitude,nrow=2)
  facet_wrap(scenario~magnitude,nrow=2)


#ggsave("paper_figure/proportion_positive.png",width=8,height=5)

prop_zone_plot

att_gross_nodate_data <- 
  bind_rows(
    att_gross_nodate_zone |> 
      filter(scenario=="SSP585") |> 
      filter(index=="value")     ,
    
    att_gross_nodate_zone |> filter(index=="quantity") |> 
      pivot_wider(names_from=region,values_from = value) |> 
      mutate(across(north:se,~./CA)) |> 
      pivot_longer(CA:se,names_to = "region") |> 
      mutate(value=value*100)) |> 
  filter(!region=="CA") |>         
  extmodify()

att_gross_zone_plot <- ggplot(
  att_gross_nodate_data |> 
    filter(index=="value") |> 
    mutate(ext_type=str_replace(ext_type,"_"," & \n")) |> 
    filter(scenario=="SSP585"), 
  aes(x=ext_type,y=value,fill=region))+
  geom_boxplot()  +
  # facet_wrap(
  #   index~.,scales = "free",nrow = 2,
  #   labeller = as_labeller(c(value ="Extreme NPP TgC/yr", 
  #                            quantity="Frequency (%)")),
  #   strip.position = "left") +
  labs(x="",y="",fill="",tag="b")+
  labs(y=expression(Extreme~NPP~TgC~yr^-1))  +
  theme_bw() +
  theme(axis.ticks.length.y = unit(-0.1,"cm"),
        axis.text.y = element_text(angle=90),
        strip.background = element_blank(),
        strip.placement = "outside")

att_gross_zone_plot
# the attribution of extreme plot

att_space <- att_nodatemodel |> 
  group_nest(ext_type) |> 
  mutate(unit=str_c("10th npp attributed to ",ext_type)) |> 
  mutate(unit=as.list(unit)) |>
  mutate(bound=list(sf::st_as_sf(climatebound))) 


att_space_plot <-pmap(
  list(att_space$data,
       att_space$bound,
       att_space$unit),
  ~ggplot()+
    geom_raster(data =..1,aes(x = x, y = y,fill = value)) +
    geom_sf(data=..2,fill=alpha(0.01)) +
    facet_wrap(scenario~.,nrow = 1)+
    coord_sf()+
    scale_fill_gradientn(colours = terrain.colors(110))+
    labs(title = ..3,
         fill=expression("npp C g"~m^-2~year^-1),
         x ='Longitude(E)',y="Latitude(N)") +
    space_theme +
    theme_bw())


#regionalized_space


prop_zone_plot/att_gross_zone_plot+
#att_space_plot[[1]]+
  plot_annotation(tag_levels = "a")

ggsave("paper_figure/prop.png",width=6,height=6,dpi=700,scale=1.1) 

```

#Result3 Extreme pattern

```{r}
# load("rdata/ca_density_mean.RData")

load("rdata/ca_gross_sum80_zonesum.RData")
#load("rdata/ca_density_sum80.RData")
```

## Negative TS & Magnitude

```{r}
#| label: extreme_ts


ext1_data <- ext_nospace_zone |> 
  filter(var=="npp"&magnitude=="tenth")  |> 
  summarise(error=sd(value),value=mean(value),.by=c(scenario,var,date,magnitude))
  

ext1_sig <- ext1_data                                 |> 
  group_nest(scenario)                                |> 
  mutate(model=map(data,~lm(data=.,value~date)        |> 
                   broom::tidy()))                    |> 
  select(-data)                                       |> 
  unnest(model)                                       |> 
  filter(term=="date")                                |> 
  mutate(x=c(2035,2035,2055,2055),y=c(-78,-84,-78,-84)) |>  # -78, 80 for the figure without ribbon
  mutate(p.value=round(p.value,digits=5))             |> 
  mutate(psymble = case_when(p.value<0.001 ~ "***",
                             p.value<0.01 ~ "**",
                             p.value<0.05  ~ "*"
                             ))    |> 
  mutate(statistic=round(statistic,digits=2))         |> 
  mutate(label=str_c(scenario,": ",statistic,psymble)) 
  


 ext1 <- ext1_data |>
  ggplot(aes(x=date,y=value)) +
  geom_text(data=ext1_sig,aes(x=x,y=y-30,label=label,col=scenario),
            family="serif",size=3,show.legend = FALSE)+
  geom_ribbon(aes(ymin = value-error,ymax =value+error,fill=scenario),alpha = 0.2,show.legend = FALSE) +
  geom_line(aes(col=scenario))+
  geom_smooth(method="lm",se=FALSE,aes(col=scenario),linewidth=0.5) +
  labs(y=expression("NPP gC"~m^-2~yr^-1),col="",x="")+
  theme_bw()+
  theme(legend.position = "top")


ext1

ext2 <- ext_nodate_zone |> 
  
  mutate(value=value*80)  |> 
  
  
  filter(index=="value"&magnitude=="tenth"&var=="npp")  |>  
  extmodify()                                          |> 
  ggplot(aes(x=region,y=value,fill=scenario,col=scenario))+
    stat_summary(
      fun.min = function(x) {mean(x) - sd(x)/sqrt(length(x))},
      fun.max = function(x) {mean(x) + sd(x)/sqrt(length(x))},
      geom = "errorbar",
      position=position_dodge(0.9),width = 0.9) +  
    stat_summary(fun=mean,geom = "bar",
      position=position_dodge(0.9),width = 0.9) +
      ylab(expression(NPP~gC~m^-2))+
      labs(fill="",col="",x="") +
      theme_bw()  +
      theme(legend.position = "top") 


ext1/ext2+
  plot_annotation(tag_levels = "a")


ggsave("paper_figure/extreme_date_magnitude.png",width=20,#width=25,
       height=20,units="cm",scale = 0.7)


```

## Positive TS & Magnitude

```{r}
ext1_data <- ext_nospace_zone |> 
  filter(var=="npp"&magnitude=="ninetieth")  |> 
  summarise(error=sd(value),value=mean(value),.by=c(scenario,var,date,magnitude))


dis1 <-218  

ext1_sig <- ext1_data                                 |> 
  group_nest(scenario)                                |> 
  mutate(model=map(data,~lm(data=.,value~date)        |> 
                   broom::tidy()))                    |> 
  select(-data)                                       |> 
  unnest(model)                                       |> 
  filter(term=="date")                                |> 
  mutate(x=c(2035,2035,2055,2055),y=c(-78+dis1,-84+dis1,-78+dis1,-84+dis1)) |>  # -78, 80 for the figure without ribbon
  mutate(p.value=round(p.value,digits=5))             |> 
  mutate(psymble = case_when(p.value<0.001 ~ "***",
                             p.value<0.01 ~ "**",
                             p.value<0.05  ~ "*"
                             ))    |> 
  mutate(statistic=round(statistic,digits=2))         |> 
  mutate(label=str_c(scenario,": ",statistic,psymble)) 
  


 ext1 <- ext1_data |>
  ggplot(aes(x=date,y=value)) +
  geom_text(data=ext1_sig,aes(x=x,y=y-30,label=label,col=scenario),
            family="serif",size=3,show.legend = FALSE)+
  geom_ribbon(aes(ymin = value-error,ymax =value+error,fill=scenario),alpha = 0.2,show.legend = FALSE) +
  geom_line(aes(col=scenario))+
  geom_smooth(method="lm",se=FALSE,aes(col=scenario),linewidth=0.5) +
  labs(y=expression("NPP gC"~m^-2~yr^-1),col="",x="")+
  theme_bw()+
  theme(legend.position = "top")
 



ext2 <- ext_nodate_zone |> 
  
  mutate(value=value*80)  |> 
  
  
  filter(index=="value"&magnitude=="ninetieth"&var=="npp")  |>  
  extmodify()                                          |> 
  ggplot(aes(x=region,y=value,fill=scenario,col=scenario))+
    stat_summary(
      fun.min = function(x) {mean(x) - sd(x)/sqrt(length(x))},
      fun.max = function(x) {mean(x) + sd(x)/sqrt(length(x))},
      geom = "errorbar",
      position=position_dodge(0.9),width = 0.9) +  
    stat_summary(fun=mean,geom = "bar",
      position=position_dodge(0.9),width = 0.9) +
      ylab(expression(NPP~gC~m^-2))+
      labs(fill="",col="",x="") +
      theme_bw()  +
      theme(legend.position = "top") 


ext1/ext2+
  plot_annotation(tag_levels = "a")
```

These data derived from gross data are summed in the region scale

Under certain situation, the number of each year for extreme value is too small to produce the values instact number of 20 larger and 20 small.

## whole

```{r}
load("rdata/ca_gross_sum80_zonesum.RData")
# Create the plot and assign it to an object
npp_magnitude <- ext_nodate_zone |> 
#  mutate(value = value * 80) |> 
  filter(index == "value", var == "npp") |>  
  extmodify() |> 
  mutate(magnitude = factor(magnitude, levels = c("ninetieth", "tenth"))) |> 
  ggplot(aes(x = region, y = value, fill = scenario, color = scenario)) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "bar", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
   facet_grid(
    magnitude ~ ., 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  labs(
    x = NULL, 
    # y = expression(NPP ~ (gC ~ m^{-2}~yr^{-1})), 
    y = expression(NPP ~ (TgC ~ yr^{-1})),
  # y = expression(NPP ~ (gC ~ yr^{-1})), 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
    theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    axis.text.y = element_text(angle = 90),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1),
    axis.ticks.length.y = unit(-0.1, "cm"),
    strip.text.y = element_text(size = rel(1.1)),
    plot.tag = element_text(size = 14, face = "bold")
  )+
    theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
  
  
npp_sum

ggsave(
  filename = "paper_figure/NPP_extremes_proport1_improved.png",
  plot = npp_plot,
  width = 4,
  height = 6.5,
  units = "in",
  dpi = 600,
  bg = "transparent" 
)



npp_quantity <- ext_nodate_zone |> 
  filter(index == "quantity", var == "npp") |>  
#  filter(region!="CA") %>% 
  extmodify() |> 
  mutate(magnitude = factor(magnitude, levels = c("ninetieth", "tenth"))) |> 
  ggplot(aes(x = region, y = value, fill = scenario, color = scenario)) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "bar", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
   facet_grid(
    magnitude ~ ., 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  labs(
    x = NULL, 
    y = "The quantity of NPP extremes",
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
    theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    axis.text.y = element_text(angle = 90),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1),
    axis.ticks.length.y = unit(-0.1, "cm"),
    strip.text.y = element_text(size = rel(1.1)),
    plot.tag = element_text(size = 14, face = "bold")
  )+
    theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
  




load("rdata/ca_density_mean.RData")



# Create the plot and assign it to an object
npp_average <- ext_nodate_zone |> 
  filter(index == "value", var == "npp") |>  
  extmodify() |> 
  mutate(magnitude = factor(magnitude, levels = c("ninetieth", "tenth"))) |> 
  ggplot(aes(x = region, y = value, fill = scenario, color = scenario)) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "bar", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
   facet_grid(
    magnitude ~ ., 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  labs(
    x = NULL, 
     y = expression(NPP ~ (gC ~ m^{-2}~yr^{-1})), 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
    theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    axis.text.y = element_text(angle = 90),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1),
    axis.ticks.length.y = unit(-0.1, "cm"),
    strip.text.y = element_text(size = rel(1.1)),
    plot.tag = element_text(size = 14, face = "bold")
  )+
    theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )

(npp_magnitude|npp_average|npp_quantity) +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides = "collect")&
  theme(legend.position='top') 


ggsave("paper_figure/NPP_extreme_series.png",width=9,height=6,dpi=700)


```

## Separate date

```{r}

load("rdata/ca_gross_sum80_zonesum.RData")

ext_nospace_date_zone |> 
  filter(var=="npp") |> 
  ggplot(aes(x = scenario, y = value, fill = class, color = class,group = class)) +
  # geom_col(position = "dodge") +
   facet_grid(
    magnitude ~. , 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
  labs(
    x = NULL, 
    y = expression(NPP ~ (Tg~C~yr^{-1})), 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    text = element_text(family = "Arial", size = 12),
    axis.text = element_text(size = 10),
    strip.text.y = element_text(angle = 0, size = 11)
  )# +
  geom_signif(comparisons=list(c("SSP126", "SSP585")),
               tip_length = 0.04, vjust=0.2,color="black")




load("rdata/ca_density_sum80.RData")
load("rdata/ca_density_mean.RData")

ext_nospace_date_zone |> 
  filter(var=="npp") |> 
  ggplot(aes(x = scenario, y = value, fill = class, color = class,group = class)) +
  # geom_col(position = "dodge") +
   facet_grid(
    magnitude ~. , 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
  labs(
    x = NULL, 
    y = expression(NPP ~ (gC ~ m^{-2}~yr^{-1})), 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    text = element_text(family = "Arial", size = 12),
    axis.text = element_text(size = 10),
    strip.text.y = element_text(angle = 0, size = 11)
  )+
  geom_signif(comparisons=list(c("SSP126", "SSP585")),
               tip_length = 0.04, vjust=0.2,color="black")


ext_nospace_date_zone |> 
  filter(var=="npp") |> 
  ggplot(aes(x = scenario, y = quantity, fill = class, color = class,group = class)) +
  # geom_col(position = "dodge") +
   facet_grid(
    magnitude ~. , 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
  labs(
    x = NULL, 
    y = "Frequency per year", 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    text = element_text(family = "Arial", size = 12),
    axis.text = element_text(size = 10),
    strip.text.y = element_text(angle = 0, size = 11)
  )+
  geom_signif(comparisons=list(c("SSP126", "SSP585")),
               tip_length = 0.04, vjust=0.2,color="black")



ext_nospace_date_zone |> 
  filter(var=="pr") |> 
  ggplot(aes(x = scenario, y = quantity, fill = class, color = class,group = class)) +
  # geom_col(position = "dodge") +
   facet_grid(
    magnitude ~. , 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) +
  labs(
    x = NULL, 
    y = "The Frequency of Precipitation per year", 
    fill = NULL, 
    color = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside",
    text = element_text(family = "Arial", size = 12),
    axis.text = element_text(size = 10),
    strip.text.y = element_text(angle = 0, size = 11)
  )+
  geom_signif(comparisons=list(c("SSP126", "SSP585")),
               tip_length = 0.04, vjust=0.2,color="black")



ext_nospace_date_zone |> 
  filter(var=="tas") |> 
  ggplot(aes(x = scenario, y = quantity, fill = class, color = class,group = class)) +
  # geom_col(position = "dodge") +
   facet_grid(
    magnitude ~. , 
    scales = "free_y", 
    switch = "y", 
    labeller = labeller(
      magnitude = c(
        "ninetieth" = "Positive extremes",
        "tenth"="Negative extremes"
      )
    )
  ) +
  stat_summary(
    fun.min = \(x) mean(x) - sd(x) / sqrt(length(x)),
    fun.max = \(x) mean(x) + sd(x) / sqrt(length(x)),
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.9
  ) +
  stat_summary(
    fun = mean, 
    geom = "point", 
    position = position_dodge(width = 0.9), 
    width = 0.9
  ) 


```

